---
title: 'homework iv'
author: 'Nirbhay Pherwani, Dhiren Chandnani'
date: '2019-10-04'
output:
    pdf_document:
    latex_engine: xelatex
header-includes:
  \usepackage{booktabs}
---

```{r echo=FALSE}
# This chunk is just to make it possible to shrink the typeface in succeeding chunks. Mainly this will be used for the crosstabs.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Initialization
Here we load the tidyverse packages and the `data.table` package and load the nyc311 data set. Then we fix the column names of the nyc311 data so that they have no spaces.

```{r initialize}
library(tidyverse)
library(data.table)
nyc311<-fread("311_Service_Requests_from_2010_to_Present.csv")
names(nyc311)<-names(nyc311) %>%
  stringr::str_replace_all("\\s", ".")
```
# Tidying the NYC311 Data

## Removing unwanted columns

```{r removeColumns}
nyc311 <- subset(nyc311, select= -c(Landmark, Park.Borough, School.City, School.State, School.Zip, Taxi.Company.Borough))
```

## Complaints
We observed that there were 243 unique complaint types. A lot of complaints had just 1 or 2 instances. So for analysis only major complaints would be required. Hence, only complaints with counts greater than 60 are selected to maintain relevancy. 

```{r complaintGreaterThan60}
nyc311<-nyc311 %>%
  group_by(Complaint.Type) %>%
  filter(n() >= 60)
```

Complaint types had unwanted characters and also were not in a formatted manner. So we made it all upper case and replaced those characters.
```{r makeUpperCaseAndReplaceComplaint}
library(dplyr)
nyc311$Complaint.Type <- toupper(nyc311$Complaint.Type)
nyc311$Complaint.Type <- gsub('/', ' ', nyc311$Complaint.Type)
nyc311$Complaint.Type <- gsub('-', ' ', nyc311$Complaint.Type)
```

## Location Types
We observed that there were 138 unique Location types. A lot of locations had just 1 or 2 instances. So for analysis only major locations would be required. Hence, only locations with counts greater than 10 are selected to maintain relevancy.

```{r locationGreaterThan10}
nyc311<-nyc311 %>%
  group_by(Location.Type) %>%
  filter(n() >= 10)
```

## Agencies
We observed that there were 64 unique Agency types. A lot of Agencies had just 1 or 2 instances. So for analysis only major agencies would be required. Hence, only agencies with counts greater than 10 are selected to maintain relevancy.

```{r AgencyGreaterThan10}
nyc311<-nyc311 %>%
  group_by(Agency) %>%
  filter(n() >= 10)
```

## Community Board
Community Board had a combination of location code and the Borough name. As we already had the Borough column, we just kept the community code using tidyr separate command.

```{r separateCommunityBoard}
nyc311<-nyc311 %>% separate(Community.Board, c("Community.Code"), sep=" ", extra = "drop")
```

## Incident Address
Incident Address had a combination of location code and the Street name. As we already had the street name column, we just kept the location code using tidyr separate command.

```{r incidentAddressChange}
nyc311<-nyc311 %>% separate(Incident.Address, c("Incident.Code"), sep=" ",extra = "drop")
```

## Zipcodes
The zipcodes were not in a correct 5-digit format. We used the zipcode package to clean the zip codes.

```{r cleanZip}
#install.packages("zipcode")
library(zipcode)
nyc311$Incident.Zip<-clean.zipcodes(nyc311$Incident.Zip)
```

## City
Although the city columnn seems a little ambiguous, we did not delete it altogether. We just changed the case to Upper for now.

```{r makeUpperCaseAndReplaceCity}
library(dplyr)
nyc311$City <- toupper(nyc311$City)
```

## Missing Values
We wanted to replace all kinds of missing values to NA, but since the data count is too high, it was turning into a very expensive operation. We will think of faster ways to achieve this. For now we have included the code (Commented) that we tried.
```{r replaceMissingValues}
#nyc311[nyc311=="" | nyc311=="N/A" | nyc311=="N / A" | nyc311==" " | nyc311=="UNKNOWN"]<-NA
```

# Second Dataset

## Importing Dataset
We have selected the Housing New York Units by Building data. The data is provided by Department of Housing Preservation and Development (HPD) and is avaliable from the NYC Open Data website (https://data.cityofnewyork.us/Housing-Development/Housing-New-York-Units-by-Building/hg8x-zxpr)

```{r getNewData}
nycHousingData <-fread("https://data.cityofnewyork.us/api/views/hg8x-zxpr/rows.csv")
names(nycHousingData)<-names(nycHousingData) %>%
  stringr::str_replace_all("\\s", ".")
nycHousingData[nycHousingData=="" | nycHousingData==" "]<-NA
```

The data is clean. Only the Borough names are not in the same case as the NYC 311 data. So we decided to change the case.

```{r makeBoroughUpperCase}
nycHousingData$Borough <- toupper(nycHousingData$Borough)
```


